### 2014-04-21 tex/main-body-text.tex
Lets try to push to my repo and then do a pull request
Add description of structure modeling and Gibbs samlping therein.
I have sort of just put that in there.  It could use some editing, no
doubt.

\section*{Comparison of methods}
We start with a succinct mathematical description of the {\sc structure} algorithm, and then\ldots

\subsection*{{\sc structure}}

>>>>

\section*{Comparison of methods}
We start with a succinct mathematical description of the {\sc structure}
algorithm, then describe the {\sc flock} algorithm, and finally explain
the close relationship between the two.

\subsection*{{\sc structure}}
In the {\sc structure} model with no admixture, the unknown ``subpopulation'' that the 
$i\thh$ individual ($i=1,\ldots,N)$ belongs to is denoted by $Z_i \in \{1,\ldots,K\}$, where $K$ is the 
number of subpopulations (or clusters, as they are often referred to).  
In a diploid individual from cluster $k$, the 
allelic types of the two gene copies at the $\ell\thh$ locus are assumed
to be drawn independently from the vector of allele frequencies at
locus $\ell$,  $\theta_{k\ell}=(\theta_{k\ell 1},\ldots,\theta_{k\ell A_\ell})$, where $A_\ell$ 
denotes the number of alleles observed in the data set at locus $\ell$.
Hence, if $Y_{ij}$ denotes a vector of length $A_\ell$ whose components denote the number
of copies of each of the $A_\ell$ alleles at locus $j$ in individual $i$, and 
$Z_i=k$, then $Y_{i\ell}$ follows the multinomial distribution of two trials with
with $A_\ell$  components and cell probabilities given by the allele frequencies: 
\begin{equation}
(Y_{i\ell}~|~Z_i=k) \sim \mathrm{Mult}_{A_\ell}(2, \theta_{k\ell}).
\end{equation}
In the model without physical linkage, the genotypes at the loci are assumed to
be independent of one another, so the probability of the genotype data at
all $L$ loci---$Y_i=(Y_{i1},\ldots,Y_{iL})$---is simply a product of multinomial probabilities.
In the uncorrelated allele frequencies model, the prior on each $\theta_{k\ell}$ is a
Dirichlet distribution with parameters $(\lambda_{k\ell1},\ldots,\lambda_{k\ell A_\ell})$,
which are usually set to a value like $1/A_\ell$.

Inference in the model proceeds by sampling from the joint posterior of the 
$Z_i$'s and $\theta = (\theta_1,\ldots,\theta_K)$ using Gibbs sampling.  That is,
at iteration $t$:
\begin{enumerate}
\item each $Z^{(t)}_i$ is updated to $Z^{(t+1)}_i$ by sampling a value from 
the full conditional distribution of $Z_i$ given
$Y_i$ and $\theta^{(t)}$ (the current estimate of the allele frequencies).  This
distribution is found using Bayes' theorem. In {\sc structure}'s formulation, the 
prior probability that $Z_i=k$ is $1/K$ for all $k=1,\ldots, K$.     
\item $\theta^{(t)}$ is updated from its full conditional distribution.  For each cluster
$k$, and locus $\ell$ the full conditional for $\theta_{k\ell}$ is independently a 
Dirichlet distribution with parameters $\lambda_{k\ell j} + \#(k,\ell,j)$, for $j=1,\ldots, A_\ell$,
where $\#(k,\ell,j)$ is the number of alleles of type $j$ at locus $\ell$ observed in 
individuals whose $Z^{(t+1)}_i = k$.   
\end{enumerate}

It will be useful for our comparison with {\sc flock} to point out that another way 
of pursuing inference in this model would be to first integrate out the 
Dirichlet priors on the allele frequencies and then sample from the posterior
for the $Z_i$'s by Gibbs sampling.  When $\theta$ is integrated out, the genotypes 
of the individuals are no longer conditionally independent (they were originally
independent {\em conditional} on $\theta$ and the $Z_i$'s), so the calculation of 
the full conditional distribution of $Z_i$ is now somewhat more involved: the full 
conditional distribution of $Y_{i\ell}$ given $Z_i=k$, and $Z_{(-i)}$ and $Y_{(-i)\ell}$---the
cluster memberships and genotypes, respectively,  of all the remaining individuals---is
now a Compound Dirichlet Multinomial distribution (CDM):
\begin{eqnarray}
\lefteqn{(Y_{i\ell}~|~Z_i=k,~Z_{(-i)},~Y_{(-i)\ell}) \sim} \label{eq:cdm}\\
& & \mathrm{CDM}(\lambda_{k\ell 1} + \#_{(-i)}(k,\ell,1), \ldots,
\lambda_{k\ell A_\ell} + \#_{(-i)}(k,\ell,A\ell)), \nonumber
\end{eqnarray}
where $\#_{(-i)}(k,\ell,1)$ is the number of alleles of type $j$ at locus $\ell$
found in individuals {\em other than individual $i$} that currently
belong to cluster $k$. Calculating (\ref{eq:cdm}) for each value of $Z_i=k \in \{1,\ldots,K\}$ 
and normalizing to sum to one gives the full conditional distribution for 
$Z_i$:
\begin{equation}
P(Z_i=k~|~Y_i, ~Z_{(-i)},~Y_{(-i)\ell})~~,~~k=1,\ldots,K.
\label{eq:fc}
\end{equation}
A new value of $Z_i$ would be drawn from (\ref{eq:fc}) if doing Gibbs sampling in this
version of the {\sc structure} model in which $\theta$ has been integrated out.  In fact,
this is the approach (with a slightly different prior weight on the $Z_i$'s) taken to update 
$Z_i$ in both {\sc hwler} \citep{Pel&Mas2006} and {\sc structurama} \cite{Hue&And2007} when not
proposing changes to the number of clusters.
